---
title: "create_results_combined_with_method"
output: html_document
---

#Using OR
```{r}
#loading required libraries
library(metafor) 
library(evd)
library(parallel)
library(repMeta)
library(Matrix)
library(metadat)
library(numDeriv)
library(pbapply)
library(pbmcapply)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(webshot2)
library(knitr)
library(dplyr)
library(purrr)


load("binary.RData")

calI.cont <- function(dat1){
  weight <- 1/dat1$s2
  theta_f <- weighted.mean(dat1$y,weight)
  Q <- sum(weight*(dat1$y-theta_f)^2)
  H2 <- Q/(nrow(dat1)-1)
  I2 <- (Q>(nrow(dat1)-1))*(1-1/H2)
  # S <- sum(weight)-sum(weight^2)/sum(weight)
  # D_l_tau <- (Q>(nrow(dat1)-1))*(Q-(nrow(dat1)-1))/S
  # # w_as <- 1/(dat1$s2+D_l_tau)
  # # theta_r <- weighted.mean(dat1$y,w_as)
  # # var.theta_r <- 1/sum(w_as)
  return(I2)
}

####finding unique meta analysis for each Cochran Dataset
ma.id.ls <- unique(binary.cochrane$ma.id)

# summarize for each meta-analysis, how much studies we have. The goal is to identify ma.id that has more than 2 studies.
#This calculates the number of rows (i.e., studies) for each ma.id using mclapply.
length.ma <- parallel::mclapply(ma.id.ls,mc.cores=6,function(ma.id){
  dat <- binary.cochrane[binary.cochrane$ma.id == ma.id,]
  return(nrow(dat))
}) 

length.ma <- unlist(length.ma)

#identifying which ma.id are greater than 2
ma.id.lsgt2 <- ma.id.ls[length.ma>2]


##test run for the first 50 rows
parallel::detectCores()
test_OR <- parallel::mclapply(ma.id.lsgt2, mc.cores=4, function(ma.id){
  tryCatch({
    dat <- binary.cochrane[binary.cochrane$ma.id == ma.id,]
    dat <- metafor::escalc(measure = "OR",
                           ai = dat$r1,
                           bi = dat$n1 - dat$r1,
                           ci = dat$r2,
                           di = dat$n2 - dat$r2,
                           data = dat)
    dat <- data.frame(y = dat$yi, s2 = dat$vi)
    Rmls <- repMeta::Rm.func.iterative(dat,1)

    return(c(Rmls$stat[1], Rmls$P_1, length(Rmls$out_studies)))

  }, error = function(e){
    return(c(NA, NA, NA))
  })
})


# diagnostic only, copy the same code within bracket of mclapply into the for loop below to find out which step gives you errors
for (ma.id in ma.id.lsgt2[1:50]){
  dat <- binary.cochrane[binary.cochrane$ma.id == ma.id,]
  # metafor::escalc(dat)
  dat <- metafor::escalc(measure = "OR", 
                         ai = dat$r1, 
                         bi = dat$n1-dat$r1, 
                         ci = dat$r2, 
                         di = dat$n2-dat$r2, 
                         data = dat)
  dat <- data.frame(
    y = dat$yi,  # effect size estimates
    s2 = dat$vi  # variances
  )
  I2 <- calI.cont(dat) # I2
  # repMeta::calR(dat)$maxR
  Rmls <- repMeta::Rm.func.iterative(dat,1)
  
  print(c(Rmls$stat[1],
    Rmls$P_1,
    length(Rmls$out_studies)))
}

##organize OR results into a dataframe
results_OR <- data.frame(
  ma.id = ma.id.lsgt2,
  Rm = sapply(test_OR, function(x) x[1]),
  pval = sapply(test_OR, function(x) x[2]),
  out_studies = sapply(test_OR, function(x) x[3])
)

head(results_OR)
write.csv(results_OR, "results_OR_all.csv", row.names = FALSE)
```

#Using RR
```{r}
##test run for all rows - RR
test_RR <- parallel::mclapply(ma.id.lsgt2, mc.cores = 4, function(ma.id){
  tryCatch({
    dat <- binary.cochrane[binary.cochrane$ma.id == ma.id,]

    dat <- metafor::escalc(
      measure = "RR",
      ai = dat$r1,
      bi = dat$n1 - dat$r1,
      ci = dat$r2,
      di = dat$n2 - dat$r2,
      data = dat
    )

    dat <- data.frame(y = dat$yi, s2 = dat$vi)

    Rmls <- repMeta::Rm.func.iterative(dat, 1)

    return(c(Rmls$stat[1], Rmls$P_1, length(Rmls$out_studies)))

  }, error = function(e){
    return(c(NA, NA, NA))
  })
})



# diagnostic only, copy the same code within bracket of mclapply into the for loop below to find out which step gives you errors
for (ma.id in ma.id.lsgt2[1:50]) {

  dat <- binary.cochrane[binary.cochrane$ma.id == ma.id,]

  dat <- metafor::escalc(
    measure = "RR",
    ai = dat$r1,
    bi = dat$n1 - dat$r1,
    ci = dat$r2,
    di = dat$n2 - dat$r2,
    data = dat
  )

  dat <- data.frame(y = dat$yi, s2 = dat$vi)

  Rmls <- repMeta::Rm.func.iterative(dat, 1)

  print(c(Rmls$stat[1], Rmls$P_1, length(Rmls$out_studies)))
}


##organize RR results into a dataframe
results_RR <- data.frame(
  ma.id = ma.id.lsgt2,
  Rm = sapply(test_RR, function(x) x[1]),
  pval = sapply(test_RR, function(x) x[2]),
  out_studies = sapply(test_RR, function(x) x[3])
)

head(results_RR)
write.csv(results_RR, "results_RR_all.csv", row.names = FALSE)
```

#Using RD
```{r}
##test run for all rows - RD
library(R.utils)

test_RD <- parallel::mclapply(
  ma.id.lsgt2, mc.cores=4,
  function(ma.id){

    tryCatch({
      withTimeout({

        dat <- binary.cochrane[binary.cochrane$ma.id == ma.id,]

        dat <- metafor::escalc(
          measure="RD",
          ai=dat$r1, bi=dat$n1-dat$r1,
          ci=dat$r2, di=dat$n2-dat$r2,
          data=dat
        )

        dat <- data.frame(y = dat$yi, s2 = dat$vi)

        Rmls <- repMeta::Rm.func.iterative(dat,1)

        return(c(Rmls$stat[1], Rmls$P_1, length(Rmls$out_studies)))

      }, timeout = 3)

    }, TimeoutException = function(e){
      return(c(NA, NA, NA))
    }, error=function(e){
      return(c(NA, NA, NA))
    })

  }
)




# diagnostic only, copy the same code within bracket of mclapply into the for loop below to find out which step gives you errors
for (ma.id in ma.id.lsgt2[1:50]) {

  dat <- binary.cochrane[binary.cochrane$ma.id == ma.id,]

  dat <- metafor::escalc(
    measure = "RD",
    ai = dat$r1,
    bi = dat$n1 - dat$r1,
    ci = dat$r2,
    di = dat$n2 - dat$r2,
    data = dat
  )

  dat <- data.frame(y = dat$yi, s2 = dat$vi)

  Rmls <- repMeta::Rm.func.iterative(dat, 1)

  print(c(Rmls$stat[1], Rmls$P_1, length(Rmls$out_studies)))
}


##organize RD results into a dataframe
results_RD <- data.frame(
  ma.id = ma.id.lsgt2,
  Rm = sapply(test_RD, function(x) x[1]),
  pval = sapply(test_RD, function(x) x[2]),
  out_studies = sapply(test_RD, function(x) x[3])
)


head(results_RD)
write.csv(results_RD, "results_RD_all.csv", row.names = FALSE)
```


#Combine 3 results - OR, RR, RD + add method column
```{r}
#transform result to wide format
results_wide <- reduce(
  list(results_OR, results_RR, results_RD),
  full_join,
  by = "ma.id"
)
results_clean <- results_wide %>%
  rename(
    Rm_OR = Rm.x,
    p_OR = pval.x,
    out_OR = out_studies.x,
    
    Rm_RR = Rm.y,
    p_RR = pval.y,
    out_RR = out_studies.y,
    
    Rm_RD = Rm,
    p_RD = pval,
    out_RD = out_studies
  )

write.csv(results_clean, "results_combined_wide.csv", row.names = FALSE)
```

#Goal：
#Automatically identify which effect measure (RD, RR, or OR) each meta-analysis (MA) in the binary.cochrane dataset was based on

```{r}
# identify which measure MA is using
# RR
 dat <- binary.cochrane[binary.cochrane$ma.id == ma.id,]
  dat <- metafor::escalc(measure = "RR", 
                         ai = dat$r1, 
                         bi = dat$n1-dat$r1, 
                         ci = dat$r2, 
                         di = dat$n2-dat$r2, 
                         data = dat)
  dat <- data.frame(
    y = dat$yi,  # effect size estimates
    s2 = dat$vi  # variances
  )
  
  rr.y <- dat$y[1]
  rr.s2 <- dat$s2[1]
  
# RD
  dat <- binary.cochrane[binary.cochrane$ma.id == ma.id,]
  dat <- metafor::escalc(measure = "RD", 
                         ai = dat$r1, 
                         bi = dat$n1-dat$r1, 
                         ci = dat$r2, 
                         di = dat$n2-dat$r2, 
                         data = dat)
  dat <- data.frame(
    y = dat$yi,  # effect size estimates
    s2 = dat$vi  # variances
  )
  
  rd.y <- dat$y[1]
  rd.s2 <- dat$s2[1]
  
# OR
  dat <- binary.cochrane[binary.cochrane$ma.id == ma.id,]
  dat <- metafor::escalc(measure = "OR",
         ai = dat$r1,
         bi = dat$n1-dat$r1,
         ci = dat$r2,
         di = dat$n2-dat$r2,
         data = dat)
  dat <- data.frame(
    y = dat$yi,  # effect size estimates
    s2 = dat$vi  # variances
  )
    or.y <- dat$y[1]
    or.s2 <- dat$s2[1]
  

  # eff.m.ls <- round(c(rd.y,rd.s2,rr.y,rr.s2,or.y,or.s2),2)
  # first.study.eff.raw <- round(binary.cochrane[binary.cochrane$ma.id == ma.id,c("y","s2")][1,],2)
  # which(eff.m.ls %in% first.study.eff.raw) # 3.4 then RR, 1,2 then RD, 5,6 then OR.
  # if (all.equal(which(eff.m.ls %in% first.study.eff.raw), c(1,2))) {eff.m <- "RD"} else if (all.equal(which(eff.m.ls %in%   first.study.eff.raw), c(3,4))) {eff.m <- "RR"}
  
  eff.m.ls <- round(c(rd.y, rd.s2, rr.y, rr.s2, or.y, or.s2), 2)
  first.study.eff.raw <- round(binary.cochrane[binary.cochrane$ma.id == ma.id, c("y","s2")][1,], 2)

  match.idx <- which(eff.m.ls %in% first.study.eff.raw)

if (isTRUE(all.equal(match.idx, c(1, 2)))) {
  eff.m <- "RD"
} else if (isTRUE(all.equal(match.idx, c(3, 4)))) {
  eff.m <- "RR"
} else if (isTRUE(all.equal(match.idx, c(5, 6)))) {
  eff.m <- "OR"
} else {
  eff.m <- NA  #no match found
}

  print(paste("ma.id", ma.id, "measure =", eff.m))
```

# Combine both exact and tolerance methods (Final version)
# - First, perform exact match
# - If no match, fall back to tolerance comparison
# - This hybrid method ensures accuracy and robustness across all MAs
```{r}
results_measure <- data.frame(ma.id = numeric(), measure = character())

for (id in unique(binary.cochrane$ma.id)) {
  dat <- binary.cochrane[binary.cochrane$ma.id == id, ]
  
  rd <- metafor::escalc(measure = "RD", ai = dat$r1, bi = dat$n1 - dat$r1,
                        ci = dat$r2, di = dat$n2 - dat$r2, data = dat)
  rr <- metafor::escalc(measure = "RR", ai = dat$r1, bi = dat$n1 - dat$r1,
                        ci = dat$r2, di = dat$n2 - dat$r2, data = dat)
  or <- metafor::escalc(measure = "OR", ai = dat$r1, bi = dat$n1 - dat$r1,
                        ci = dat$r2, di = dat$n2 - dat$r2, data = dat)
  
  eff.m.ls <- round(c(rd$yi[1], rd$vi[1], rr$yi[1], rr$vi[1], or$yi[1], or$vi[1]), 4)
  first.study.eff.raw <- round(unname(as.numeric(dat[1, c("y", "s2")])), 4)
  
  idx <- which(sapply(1:(length(eff.m.ls) - length(first.study.eff.raw) + 1),
                      function(i) all(eff.m.ls[i:(i + length(first.study.eff.raw) - 1)] == first.study.eff.raw)))
  matches <- unlist(lapply(idx, function(i) i:(i + length(first.study.eff.raw) - 1)))
  
  if (all(matches == c(1, 2))) eff.m <- "RD"
  else if (all(matches == c(3, 4))) eff.m <- "RR"
  else if (all(matches == c(5, 6))) eff.m <- "OR"
  else {
    tol <- 0.001
    pairs <- list(c(1, 2), c(3, 4), c(5, 6))  #RD, RR, OR
    diffs <- sapply(pairs, function(idx) {
      all((abs(eff.m.ls[idx] - first.study.eff.raw) / (abs(first.study.eff.raw) + 1e-8)) < tol)
    })
    
    if (diffs[1]) {
      eff.m <- "RD"
    } else if (diffs[2]) {
      eff.m <- "RR"
    } else if (diffs[3]) {
      eff.m <- "OR"
    } else if (sum(diffs) > 1) {
      eff.m <- "Ambiguous"
    } else {
      eff.m <- NA
    }
  }
  
  results_measure <- rbind(results_measure, data.frame(ma.id = id, measure = eff.m))
}

write.csv(results_measure, "measure_identification_combined.csv", row.names = FALSE)
table(results_measure$measure, useNA = "ifany")
```

#Combine 3 results (OR, RR, RD) + add method column
```{r}
#Combine 3 results (OR, RR, RD) + add method column
results_OR <- read.csv("results_OR_all.csv")
results_RR <- read.csv("results_RR_all.csv")
results_RD <- read.csv("results_RD_all.csv")

# Combine
results_wide <- reduce(list(results_OR, results_RR, results_RD),
                       full_join, by = "ma.id")

# rename
results_clean <- results_wide %>%
  rename(
    Rm_OR = Rm.x,
    pval_OR = pval.x,
    out_OR = out_studies.x,
    
    Rm_RR = Rm.y,
    pval_RR = pval.y,
    out_RR = out_studies.y,
    
    Rm_RD = Rm,
    pval_RD = pval,
    out_RD = out_studies
  )

# Read measure identification
measure_info <- read.csv("measure_identification_combined.csv")

# Combine with method
results_final <- merge(results_clean, measure_info, by = "ma.id", all.x = TRUE)

# Output with clean names
write.csv(results_final,
          "results_combined_with_method.csv",
          row.names = FALSE)
```
